A linearization procedure for constrained multibody systems

### 2012-06-21 PetersonGedeHubbard2012.tex
changed table 2, fixed typos, add O'Reilly reference.

\section{Example}
\label{example}
We apply this linearization technique to the well studied rolling disc.  Figure
XX illustrates the choice of coordinates, speeds, and parameters.  The disc is
assumed to be knife edged of radius $r$, mass $m$, spin inertia $J$, and
transverse inertia $I$.  To illustrate the concept of dependent quantities, we
purposefully choose to describe the configuration and velocity of the disc with
a non-minimal set of coordinates and speeds.  This choice is not purely
academic; a physical disc equipped with a rate gyroscope and accelerometer
measure these body fixed quantities (or their time derivatives) directly,
rather than the more traditional set which is chosen for analytic convenience.
In more complicated systems with constraints, it is difficult (if not
impossible) to describe the system with a minimal choice of coordinates and
speeds; the ideas demonstrated by this simply example apply analogously to more
complicated ones.

The orientation of the disc fixed frame $C$ relative to the inertial frame $N$
is given by a sequence of angular rotations (Euler ZXY) by angles $q_0$ (yaw),
$q_1$ (lean), $q_2$ (spin).  The angular velocity of $C$ in $N$, the location
of disc center $C^*$ relative to the inertial origin $N^*$, and the velocity of
$C^*$ in $N$ are
\begin{align}
  \bm{\omega}^C &= u_0 \bm{c}_x + u_1 \bm{c}_y + u_2 \bm{c}_z \\
  \bm{r}^{C^*} &= q_3 \bm{n}_x + q_4 \bm{n}_y + q_5 \bm{n}_z \\
  \bm{v}^{C^*} &= u_3 \bm{c}_x + u_4 \bm{c}_y + u_5 \bm{c}_z
\end{align}
If the disc is to make contact with the ground, it should be obvious to the
reader that there is a configuration constraint involving the height $q_5$ of
the disc center and the lean $q_1$.  Similarly, if the disc rolls without
slip, it should be obvious that there are constraints which relate the
generalized speeds $u_i, (i = 0,\dots,5)$.  The derivation of these
constraints and the equations of motion are presented in the electronic
supplementary material and can be found in the literature (reference O'Reilly
and others?).
\begin{table}[htbp]
  \centering
  \caption{Rolling disc governing definitions and equations}
  \label{table:rollingdisc}
  \begin{tabular}[c]{l l l}
    Quantity & Shape & Description\\
    \hline
    $\bm{q},\bm{\dot{q}}$ & $\mathbf{R}^6$ & Coordinates and their time
    derivatives\\
    $\bm{u}, \bm{\dot{u}}$ & $\mathbf{R}^6$ & Speeds and their time derivatives
  \end{tabular}
  \begin{tabular}[c]{r @{ $=$ } l l}
    \multicolumn{3}{c}{ } \\
    \multicolumn{2}{c}{Equation} &           \\
    \hline
    $rc_1 + q_5$ & $0$ & $\bm{f}_c = \bm{0}$ \\
    $\left[\begin{array}{c}
        r u_1 c_2 + u_3 \\
        -r (u_0 c_2 + u_2 s_2) + u_4 \\
        r u_1 s_2 + u_5
    \end{array}\right] $ &
    $\left[\begin{array}{c}$0$\\$0$\\$0$\end{array}\right]$
      & $\bm{f}_v = \bm{0}$\\
    $\left[\begin{array}{c}
        -r (u_1 \dot{q}_2 s_2 + \dot{u}_1 c_2) + \dot{u}_3 \\
        -r ((\dot{u}_2 - u_0 \dot{q}_2) s_2   + (\dot{u}_0 + u_2 \dot{q}_2)c_2) + \dot{u}_4 \\
      r (u_1 \dot{q}_2 c_2 + s_2 \dot{u}_1) + \dot{u}_5
      \end{array}\right]$ &
      $\left[\begin{array}{c}$0$\\$0$\\$0$\end{array}\right]$ & $\bm{f}_a =
        \bm{0}$\\
        $\left[\begin{array}{c}
-\dot{q}_0 s_2 c_1 + \dot{q}_1 c_2 \\
\dot{q}_0 s_1 + \dot{q}_2 \\
\dot{q}_0 c_1 c_2 + \dot{q}_1 s_2 \\
\dot{q}_3 (-s_0 s_1 s_2 + c_0 c_2)
+ \dot{q}_4 (s_0 c_2 + s_1 s_2 c_0)
- \dot{q}_5 s_2 c_1\\
- \dot{q}_3 s_0 c_1 + \dot{q}_4 c_0 c_1 + \dot{q}_5 s_1 \\
\dot{q}_3 (s_0s_1c_2 + s_2c_0) + \dot{q}_4 (s_0s_2 - s_1c_0c_2) + \dot{q}_5 c_1 c_2
         \end{array}\right] + \left[\begin{array}{c}-u_0 \\ -u_1 \\ -u_2 \\
             -u_3 \\ -u_4 \\ -u_5 \end{array}\right]$ &
             $\left[\begin{array}{c} 0 \\ 0 \\ 0 \\ 0 \\ 0 \\ 0 \end{array}\right]$
                 & $\bm{f}_0 + \bm{f}_1 = \bm{0}$\\
                 $\left[\begin{array}{c}
-I \dot{u}_0 - m r c_2 \dot{u}_4 \\
-J \dot{u}_1 + m r s_2 \dot{u}_5 + m r c_2 \dot{u}_3 \\
-I \dot{u}_2 - m r s_2 \dot{u}_4
                   \end{array}\right]
+
                 \left[\begin{array}{c}
-I u_1 u_2 + J u_1 u_2 + g m r s_1 c_2 - m r (-u_0 u_5 + u_2 u_3) c_2 \\
                m r (u_0 u_4 - u_1 u_3) s_2 + m r (u_1 u_5 - u_2 u_4) c_2 \\
 I u_0 u_1 - J u_0 u_1 + g m r s_1 s_2 - m r (-u_0 u_5 + u_2 u_3) s_2
                   \end{array}\right]$ &
             $\left[\begin{array}{c} 0 \\ 0 \\ 0 \end{array}\right]$
               & $\bm{f}_2 + \bm{f}_3 = \bm{0}$
  \end{tabular}
\end{table}

@@

By making use of equations (\ref{eq:delta_q}) and (\ref{eq:delta_u}), we can
rewrite equation (\ref{eq:state_space_unconstrained}) as
%\begin{align}
%  \label{eq:state_space_constrained}
%  \left[
%    \begin{array}{cc}
%      \tilde{M}_{qq} & \bm{0}_{n \times o} \\
%      \tilde{M}_{uqc} & \tilde{M}_{uuc} \\
%      \tilde{M}_{uqd} & \tilde{M}_{uud}
%    \end{array}
%    \right]
%    \left[
%      \begin{array}{c}
%        \delta \bm{\dot{q}} \\
%        \delta \bm{\dot{u}}
%      \end{array}
%    \right]
%   &=
%   \left[
%     \begin{array}{cc}
%       (\tilde{A}_{qq} + \tilde{A}_{qu} C_1 ) C_0 & \tilde{A}_{qu} C_2 \\
%       (\tilde{A}_{uqc} + \tilde{A}_{uuc} C_1 ) C_0 & \tilde{A}_{uuc} C_2\\
%       (\tilde{A}_{uqd} + \tilde{A}_{uud} C_1 ) C_0 & \tilde{A}_{uud} C_2
%     \end{array}
%   \right]
%    \left[
%      \begin{array}{c}
%        \delta \bm{q}_i \\
%        \delta \bm{u}_i
%      \end{array}
%    \right]
%    +
%    \left[
%      \begin{array}{c}
%        \bm{0}_{(n+m) \times s} \\
%        \tilde{B}_{u}
%      \end{array}
%    \right]
%    \delta \bm{r}
%\end{align}
Equation \ref{eq:state_space_constrained} is a linear system of equations which
relates the time derivatives of all states (dependent and independent) to the
independent states and exogenous inputs; this is the precisely the form of
equation \ref{eq:structure}.

>>>>

\section{Example}
\label{example}
We apply this linearization technique to the well studied rolling disc.  Figure
XX illustrates the choice of coordinates, speeds, and parameters.  The disc is
assumed to be knife edged of radius $r$, mass $m$, spin inertia $J$, and
transverse inertia $I$.  To illustrate the concept of dependent quantities, we
purposefully choose to describe the configuration and velocity of the disc with
a non-minimal set of coordinates and speeds.  This choice is not purely
academic; a physical disc equipped with a rate gyroscope and accelerometer
measure these body fixed quantities (or their time derivatives) directly,
rather than the more traditional set which is chosen for analytic convenience.
In more complicated systems with constraints, it is difficult (if not
impossible) to describe the system with a minimal choice of coordinates and
speeds; the ideas demonstrated by this simple example can be applied to more
complicated ones.

The orientation of the disc fixed frame $C$ relative to the inertial frame $N$
is given by a sequence of angular rotations (Euler ZXY) by angles $q_0$ (yaw),
$q_1$ (lean), $q_2$ (spin).  The angular velocity of $C$ in $N$, the location
of disc center $C^*$ relative to the inertial origin $N^*$, and the velocity of
$C^*$ in $N$ are
\begin{align}
  \bm{\omega}^C &= u_0 \bm{c}_x + u_1 \bm{c}_y + u_2 \bm{c}_z \\
  \bm{r}^{C^*} &= q_3 \bm{n}_x + q_4 \bm{n}_y + q_5 \bm{n}_z \\
  \bm{v}^{C^*} &= u_3 \bm{c}_x + u_4 \bm{c}_y + u_5 \bm{c}_z
\end{align}
If the disc is to make contact with the ground, it should be obvious to the
reader that there is a configuration constraint involving the height $q_5$ of
the disc center and the lean $q_1$.  Similarly, if the disc rolls without
slip, it should be obvious that there are constraints which relate the
generalized speeds $u_i, (i = 0,\dots,5)$.  The derivation of these
constraints and the equations of motion are presented in the electronic
supplementary material and can be found in the literature \cite{O'Reilly1996}.
\begin{table}[htbp]
  \centering
  \caption{Rolling disc governing definitions and equations}
  \label{table:rollingdisc}
  \begin{tabular}[c]{l l l}
    Quantity & Shape & Description\\
    \hline
    $\bm{q},\bm{\dot{q}}$ & $\mathbf{R}^6$ & Coordinates and their time
    derivatives\\
    $\bm{u}, \bm{\dot{u}}$ & $\mathbf{R}^6$ & Speeds and their time derivatives
  \end{tabular}
  \begin{tabular}[c]{r @{ $=$ } l}
%    \multicolumn{3}{c}{ } \\
%    \multicolumn{2}{c}{Equation} &           \\
    \hline
    % Configuration constraint
    $\bm{f}_c$ & $rc_1 + q_5$ \\
    % Velocity constraints
    $\bm{f}_v$ & $\left[\begin{array}{c}
        r u_1 c_2 + u_3 \\
        -r (u_0 c_2 + u_2 s_2) + u_4 \\
        r u_1 s_2 + u_5
    \end{array}\right] $ \\
    % Acceleration constraints
    $\bm{f}_a$ & $\left[\begin{array}{c}
        -r (u_1 \dot{q}_2 s_2 + \dot{u}_1 c_2) + \dot{u}_3 \\
        -r ((\dot{u}_2 - u_0 \dot{q}_2) s_2   + (\dot{u}_0 + u_2 \dot{q}_2)c_2) + \dot{u}_4 \\
      r (u_1 \dot{q}_2 c_2 + s_2 \dot{u}_1) + \dot{u}_5
      \end{array}\right]$ \\
    % Kinematic differential equations
    $\bm{f}_0$ &
        $\left[\begin{array}{c}
-\dot{q}_0 s_2 c_1 + \dot{q}_1 c_2 \\
\dot{q}_0 s_1 + \dot{q}_2 \\
\dot{q}_0 c_1 c_2 + \dot{q}_1 s_2 \\
\dot{q}_3 (-s_0 s_1 s_2 + c_0 c_2)
+ \dot{q}_4 (s_0 c_2 + s_1 s_2 c_0)
- \dot{q}_5 s_2 c_1\\
- \dot{q}_3 s_0 c_1 + \dot{q}_4 c_0 c_1 + \dot{q}_5 s_1 \\
\dot{q}_3 (s_0s_1c_2 + s_2c_0) + \dot{q}_4 (s_0s_2 - s_1c_0c_2) + \dot{q}_5 c_1 c_2
         \end{array}\right]$ \\
    $\bm{f}_1$ & $\left[\begin{array}{c}-u_0 \\ -u_1 \\ -u_2 \\
             -u_3 \\ -u_4 \\ -u_5 \end{array}\right]$ \\
    % Dynamic differential equations
  $\bm{f}_2$ & $\left[\begin{array}{c}
-I \dot{u}_0 - m r c_2 \dot{u}_4 \\
-J \dot{u}_1 + m r s_2 \dot{u}_5 + m r c_2 \dot{u}_3 \\
-I \dot{u}_2 - m r s_2 \dot{u}_4
                   \end{array}\right]$ \\

  $\bm{f}_3$ & $\left[\begin{array}{c}
-I u_1 u_2 + J u_1 u_2 + g m r s_1 c_2 - m r (-u_0 u_5 + u_2 u_3) c_2 \\
                m r (u_0 u_4 - u_1 u_3) s_2 + m r (u_1 u_5 - u_2 u_4) c_2 \\
 I u_0 u_1 - J u_0 u_1 + g m r s_1 s_2 - m r (-u_0 u_5 + u_2 u_3) s_2
                   \end{array}\right]$
  \end{tabular}
\end{table}

@@

By making use of equations (\ref{eq:delta_q}) and (\ref{eq:delta_u}), we can
rewrite equation (\ref{eq:state_space_unconstrained}).  This gives Equation
\ref{eq:state_space_constrained}, a linear system of equations which relates
the time derivatives of all states (dependent and independent) to the
independent states and exogenous inputs.
